---
title: "rgee Workshop"
author: "Gopal Penny"
date: "10/25/2020"
output: 
  html_document:
    toc: true
    number_sections: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# `rgee` Workshop

The examples and most of the code below are from [https://github.com/r-spatial/rgee](https://github.com/r-spatial/rgee).

To begin we'll load `reticulate`, set our python environment, and load and initialize `rgee`.

```{r cars}
library(reticulate)
use_condaenv("r-reticulate")
library(rgee)

## It is necessary just once
# ee_install()

# Initialize Earth Engine!
ee_Initialize(email = "gopalpenny@gmail.com", drive=TRUE)

ee_check() # Check non-R dependencies
```

## Example 1: Night time lights trend

```{r}
# Function to create an image band that stores time
createTimeBand <-function(img) {
  year <- ee$Date(img$get('system:time_start'))$get('year')$subtract(1991L)
  ee$Image(year)$byte()$addBands(img)
}

# Load Night time lights image collection
collection <- ee$
  ImageCollection('NOAA/DMSP-OLS/NIGHTTIME_LIGHTS')$
  select('stable_lights')$
  map(createTimeBand) # this map command adds the time band to each image within the collection
```

Now, calculate a linear trend in each pixel

```{r}
# Calculate the trend using a Reducer, which collapses multiple images to a single image
col_reduce <- collection$reduce(ee$Reducer$linearFit())
col_reduce <- col_reduce$addBands(
  col_reduce$select('scale')) # scale is the slope of the trend

# Print metadata for the resulting image
ee_print(col_reduce)
```

Map the trend

```{r}
# Map the temporal trend. Green -- no trend. Red band -- increase, Blue band - decrease
Map$setCenter(9.08203, 47.39835, 3)
m1 <- Map$addLayer(
  eeObject = col_reduce,
  visParams = list(
    bands = c("scale", "offset", "scale"),
    min = 0,
    max = c(0.18, 20, -0.18)
  ),
  name = "stable lights trend"
)
m2 <- Map$addLayer(
  eeObject = col_reduce,
  visParams = list(
    bands = c("scale", "offset", "scale"),
    min = 0,
    max = c(0.18, 20, -0.18)
  ),
  name = "stable lights trend"
)

m1 + m2
```


## Example 2: Extract precipitation values

Load `tidyverse` for easy data wrangling and `sf` for spatial (i.e., vector) processing.

```{r}
library(tidyverse)
library(sf)

nc <- st_read(system.file("shape/nc.shp", package = "sf"), quiet = TRUE)
# nc
ggplot() + geom_sf(data=nc,aes())
```

Define the image collection containing monthly precipitation values. Note that we filter the data to the year 2001.

```{r}
terraclimate <- ee$ImageCollection("IDAHO_EPSCOR/TERRACLIMATE")$
  filterDate("2001-01-01", "2002-01-01")$
  map(function(x) x$reproject("EPSG:4326")$select("pr"))
```

Extract and map precipitation values

```{r}
# Extract precip from earth engine
ee_nc_rain <- ee_extract(x = terraclimate, y = nc, sf = FALSE)
colnames(ee_nc_rain) <- paste0("month_",sprintf("%02d", 1:12))
nc_rain <- nc %>% bind_cols(ee_nc_rain)

# Convert vector to "long" format -- i.e., 1 row per county-month combination
nc_rain_long <- nc_rain %>% gather("month","pr",starts_with("month_"))

# Use ggplot with geom_sf for spatial maps, and use facet_wrap to include a panel for each month
ggplot() + geom_sf(data=nc_rain_long,aes(fill=pr)) + facet_wrap(~month,ncol = 3)
```

Plot the same data as a timeseries

```{r}
ggplot(nc_rain_long, aes(x = month, y = pr, group = NAME, color = pr)) +
  geom_line(alpha = 0.4) +
  xlab("Month") +
  ylab("Precipitation (mm)") +
  theme_minimal()
```

## Example 3: Create an NDVI-animation

Javascript version: https://developers.google.com/earth-engine/tutorials/community/modis-ndvi-time-series-animation

```{r}
# Define the regional bounds of animation frames and a mask to clip the NDVI data by.
mask <- system.file("shp/arequipa.shp", package = "rgee") %>% 
  st_read(quiet = TRUE) %>% 
  sf_as_ee()
region <- mask$geometry()$bounds()

# Retrieve the MODIS Terra Vegetation Indices 16-Day Global 1km dataset as an ee.ImageCollection and select the NDVI band.
col <- ee$ImageCollection('MODIS/006/MOD13A2')$select('NDVI')

ee_print(mask)

# Group images by composite date
col <- col$map(function(img) {
  doy <- ee$Date(img$get('system:time_start'))$getRelative('day', 'year')
  img$set('doy', doy)
})
distinctDOY <- col$filterDate('2013-01-01', '2014-01-01')
```




```{r}
filter <- ee$Filter$equals(leftField = 'doy', rightField = 'doy');

# Define a join; convert the resulting FeatureCollection to an ImageCollection.

join <- ee$Join$saveAll('doy_matches')
joinCol <- ee$ImageCollection(join$apply(distinctDOY, col, filter))

comp <- joinCol$map(function(img) {
  doyCol = ee$ImageCollection$fromImages(
    img$get('doy_matches')
  )
  doyCol$reduce(ee$Reducer$median())
})
```


```{r}
# Create the visualization
visParams = list(
  min = 0.0,
  max = 9000.0,
  bands = "NDVI_median",
  palette = c(
    'FFFFFF', 'CE7E45', 'DF923D', 'F1B555', 'FCD163', '99B718', '74A901',
    '66A000', '529400', '3E8601', '207401', '056201', '004C00', '023B01',
    '012E01', '011D01', '011301'
    )
)

rgbVis <- comp$map(function(img) {
  do.call(img$visualize, visParams) %>% 
    ee$Image$clip(mask)
})
```


```{r}
# Create the GIF
gifParams <- list(
  region = region,
  dimensions = 600,
  crs = 'EPSG:3857',
  framesPerSecond = 10
)


print(rgbVis$getVideoThumbURL(gifParams))
```

The GIF image is located at the URL given above. The following command automatically navigates your web browser to this URL.

```{r}
browseURL(rgbVis$getVideoThumbURL(gifParams))
```

## Final thoughts

The `rgee` package is fantastic for simple data retrieval. In my experience, it has a few drawbacks compared with the native GEE javascript portal:

* Less interactive -- for example, GEE portal allows you to query raster pixels in the map. 
* `rgee` commands hold your R session hostage -- GEE doesn't process in the background
* Difficulties interrupting GEE -- may have to force quit R / RStudio
